|

### 2019-12-25 appendix/appendix.Rmd
small changes
add letter, edit appendix

The tables below show the results of the balance tests we conducted for
interventions 1 and 2. Our list of covariates are:

@@

In intervention 1, we randomised the treatment at the municipality level. We
selected 150 municipalities for the control group while the rest were placed in
the treatment group. The control group worked by deleting all the construction
sites in control assigned cities. No difference between control and treatment
variables is statistically significant.

@@

In intervention 2, we randomised the treatment at the school level. We placed
15% of schools that could potentially appear in the app in the control group.
We used block randomization, blocking by three characteristics:

@@

3. The app sends a message to the mayor's office asking details about the
	 project. If necessary, the message is also forwarded to other government
	 oversight institutions such as the city council and the Brazilian Federal
	 Comptroller. Below, we discuss the cases in which we forwarded the denounce
	 to the authorities.

@@

3.2. After we had the EWB report, we notified the mayor's office in compliance
with Brazil's Freedom of Information Act. We communicated that a citizen had
anonymously made a denounce about a school (fully identifying the school
construction that we were referring to), and that the Engineers Without Borders
found the construction delayed by a percentage that EWB provided. We submitted
a request for the mayor's office to provide an explanation on why the
construction was delayed, and asked for updated statistics on when the
construction will be finished. According to the FOI law, they had twenty days
to respond.

3.3. If the response was found insufficient, we forwarded the denounce to
the city council and to the Brazilian Federal Comptroller (CGU). As the
denounce goes to these institutions, they can harm mayors by restricting access
to Federal funds.

@@

In step (4), we relied on administrative data by the Brazilian Ministry of
Education. The Ministry of Education provides monthly data on school
constructions build with Federal Government resources and shows how many of
them are delayed across the country. We use the finishing dates reported by the
municipality at the time they sign the agreement with the Federal Government as
our measure of construction delays. This data helped us to keep the app updated
and the outcomes in the paper are selected to be six months before
(pre-treatment) and after the intervention. This data was constantly updated by
the Ministry of Education, and we found no red flags in the overall quality of
the data.

@@

In step (2), we applied the treatment by removing schools within municipalities
out of the app. In the first intervention, which happened at the municipal
level, we excluded all schools within a given randomly selected municipality.
In the second intervention, we excluded randomly selected schools out of the
app. Removing a school means that the school was in the dataset provided by the
Brazilian Ministry of Education, but it did not show up in the app. Therefore,
our treatment effect represents the school being present (treatment) or absent
(control) in the app.

@@

All these variables come from our manipulation of the Ministry of Education
data. Collecting the outcomes in this way is superior than the alternative of
gathering the data ourselves. First, we would not be able to manipulate the
results and cherry-pick favourable results. Even if they have the best
intentions, researchers may bias the results just by the choices they make in
the data collection process. For instance, if we send a RA to collect the data,
as she knows the app, she could unconsciously have a tendency to over-report
positive results. This would severely bias our findings and undermine the
credibility of our experiment. Second, using external data makes the results
easily reproducible by other researchers. As the Ministry of Education has
standardised data collection procedures, they greatly simplify and facilitate
the reproduction of our findings.

@@

We have chosen Facebook because the platform offers an easy and cheap method to
spread information. Moreover, according to the Statista website, Facebook is
used by 70 million Brazilians (60% of the entire adult population), and the
demographics largely match the demographics publicised by the Brazilian
Institute of Geography and Statistics. For instance, see the following
webpages:

@@

The code below replicates the findings of table 1 in the manuscript.

@@

	- Bjorkman, Martina and Jakob Svensson. 2009. “Power to the People: Evidence
		from a Randomized Field Experiment of a Community-Based Monitoring Project
		in Uganda.” Quarterly Journal of Economics 124(2):735–769.
	- Bjorkman, Martina and Jakob Svensson. 2010. “When Is Community-Based
		Monitoring Effective? Evidence from a Randomized Experiment in Primary
		Health in Uganda.” Journal of the European Economic Association
		8(2-3):571–581. Bjorkman Nyqvist, Martina, Damien de Walque and Jakob
		Svensson. 2017. “Experimental Evidence on the Long-Run Impact of
		Community-Based Monitoring.” American Economic Journal: Applied Economics
		9(1):33–69.

@@

* Banerjee, A. V., Banerji, R., Duflo, E., Glennerster, R., and Khemani, S.
	2010. Pitfalls of Participatory Programs: Evidence from a Randomized
		 Evaluation in Education in India. American Economic Journal: Economic
		 Policy, 2(1):1–30.
* Lieberman, E. S., Posner, D. N., and Tsai, L. L. 2014. Does Information Lead
	to More Active Citizenship? Evidence from an Education Intervention in Rural
	Kenya. World Development, 60:69–83. Raffler, Pia, Dan Posner, and Doug
	Parkerson. 2019. The Weakness of Bottom-up Accountability: Experimental
	evidence from the Ugandan health sector. URL:
	[`https://bit.ly/2L36vzI`](http://egap.org/sites/default/files/Posner%2C%20Dan_Paper.pdf)

This project re-evaluates this theoretical controversy and proposes a new field
experiment to test whether bottom-up accountability is effective to promote
better policy outcomes. The results are substantively important as well. First,
we conduct this experiment in a middle-income country, Brazil, but which has
municipalities whose social indicators match both those of developed and
developing nations. Brazil is notably unequal in several characteristics, thus
allowing us to infer results for different subgroups if required. Second, we
are evaluating school constructions, and the presence of schools have essential
consequences in terms of long-run economic development, the household income
composition, and woman empowerment.

@@

Interventions’ dates: We had three interventions here:

@@

**Assignment procedure**: The treatment here was assigned by the municipality
in the first intervention, and at the school level in the second and third
interventions.

@@

And then we checked whether the randomisation had one of the covariates
selected here significantly predicting the random assignment. We presented a
vector of acceptable random assignment in these grounds, and the Brazilian
Transparency selected the one we are using in the first intervention.

@@

1. Log of Municipal Population in 2015
2. Log of Number of Poor Families (2010 IBGE Census)
3. Log of Total Federal Transfers to the Municipality in 2016
4. IDEB Indicator for primary school quality (2015 Ministry of Education)
5. IDEB Indicator for secondary school quality (2015 Ministry of Education)

All the pre-treatment variables were not significant, as displayed in Section
\ref{sec:bal-tests}.

@@

* **Treatment groups**:
	- *Intervention 1*: Municipality with all schools funded by the Ministry of
		Education showing up in the app.
	- *Intervention 2*: School construction showing up in the app.

* **Control groups**:
	- *Intervention 1*: Municipality with all schools funded by the Ministry of
		Education **not** showing up in the app. 
	- *Intervention 2*: Selected school constructions **not** showing up in the
		app.

* **Method of manipulation delivery**:
	- *Intervention 1*: The municipalities in the treatment had all their school
		constructions showing up in the app. 
	- *Intervention 2*: The schools in the treatment group were showing up in the
		app.

@@

1. Percentage of the investment executed before the impact evaluation started
	 (placebo)
2. Percentage of the investment executed by the end of the impact evaluation
	 period
3. The difference of the percentage invested in the end and in the beginning of
	 the impact evaluation period
4. Indicator for a construction finished during the impact evaluation period
5. Indicator for a construction canceled during the impact evaluation period
6. Indicator for an updated conclusion date for the construction during the
	 impact evaluation period

@@

Below follows the CONSORT chartflow:

@@

All models use municipal-level cluster robust standard errors and state-level fixed effects. For all models, we also run regressions without clustering and without fixed effects.

# Other information

@@

* **Pre-registration**: this research has been pre-registered on the EGAP
	pre-registry tool: <https://egap.org/registration/4505>.

* **Replication materials**: The replication materials for this project are
	available here: <https://github.com/umbertomig/tdp-accountability>.

@@

* **Conflict of interests**: The authors of this analysis declare that there
	are no conflict of interest.

@@

We used `r R.version.string` to write the manuscript and this appendix. For the
regression models estimation, we used the `lfe` package. We employed the
`MatchIt` package to improve covariate matching. Everything in this report is
fully automated and can be reproduced using R Markdown.

@@

# Bibliography

>>>>

The tables below show the results of the balance tests we conducted for
interventions 1 and 2. Our covariates are:

@@

In intervention 1, we randomised the treatment at the municipality level. We
selected 150 municipalities for the control group while the rest were placed in
the treatment group. We created the control group by deleting all the
construction sites in the selected cities. No difference between control and
treatment variables is statistically significant.

@@

In intervention 2, we randomised the treatment at the school level. We placed
15% of schools that could potentially appear in the app in the control group.
We used block randomisation, blocking by three characteristics:

@@

3. The app sends a message to the mayor's office requesting details about the
	 project. If necessary, the message is also forwarded to other government
	 oversight institutions such as the city council and the Brazilian Federal
	 Comptroller (CGU). Below, we discuss the cases in which we forwarded the
	 denounce to the authorities.

@@

3.2. After we had the EWB report, we notified the mayor's office in compliance
with Brazil's Freedom of Information Act (FOI). We communicated that a citizen
had anonymously made a denounce about a school (fully identifying the school
construction that we were referring to), and that the Engineers without Borders
found the construction delayed. We submitted a request for the mayor's office
to provide an explanation on why the construction was delayed, and asked for
updated statistics on when the construction will be finished. According to the
FOI law, the mayor had twenty days to respond.

3.3. If the response was deemed insufficient, we forwarded the denounce to
the city council and to the Brazilian Federal Comptroller (CGU). As the
denounce goes to these institutions, they can harm mayors by restricting access
to Federal funds.

@@

In step (4), we relied on administrative data by the Brazilian Ministry of
Education. The Ministry of Education provides monthly data on school
constructions build with Federal Government resources and shows how many of
them are delayed across the country. We use the finishing dates reported by the
municipality at the time they sign the agreement with the Federal Government as
our measure of construction delays. This data helped us to keep the app
updated. We selected outcomes six months before and after the treatment. The
data were constantly updated by the Ministry of Education, and we found no red
flags in the overall quality of their information.

@@

In step (2), we applied the treatment by removing schools within municipalities
out of the app. In the first intervention, which happened at the municipal
level, we excluded all schools within a given randomly selected municipality.
In the second intervention, we randomly excluded selected schools out of the
app. Removing a school means that the school was in the dataset provided by the
Brazilian Ministry of Education, but it did not show up in the app. Therefore,
our treatment effect represents the school being present (treatment) or absent
(control) in the app.

@@

All these variables come from our manipulation of the Ministry of Education
data. Collecting the outcomes in this way is superior than the alternative of
gathering the data ourselves. First, we would not be able to manipulate the
results and cherry-pick favourable results. Even if they have the best
intentions, researchers may bias the results just by the choices they make in
the data collection process. For instance, if we send an RA to collect the
data, as she knows the app, she could unconsciously over-report positive
results. This would bias our findings and undermine the credibility of our
experiment. Second, using external data makes the results easily reproducible
by other researchers. As the Ministry of Education has standardised data
collection procedures, they greatly simplify and facilitate the reproduction of
our findings.

@@

We have chosen Facebook because the platform offers an easy and cheap method to
spread information. Moreover, according to the Statista website, Facebook is
used by 70 million Brazilians (60% of the entire adult population), and the
demographics largely match the demographics publicised by the Brazilian
Institute of Geography and Statistics. For instance, see the following
websites:

@@

The code below replicates table 1 in the manuscript.

@@

	- Bjorkman, Martina and Jakob Svensson. 2009. "Power to the People: Evidence
		from a Randomized Field Experiment of a Community-Based Monitoring Project
		in Uganda." Quarterly Journal of Economics 124(2):735–769.
	- Bjorkman, Martina and Jakob Svensson. 2010. "When Is Community-Based
		Monitoring Effective? Evidence from a Randomized Experiment in Primary
		Health in Uganda." Journal of the European Economic Association
		8(2-3):571–581. Bjorkman Nyqvist, Martina, Damien de Walque and Jakob
		Svensson. 2017. "Experimental Evidence on the Long-Run Impact of
		Community-Based Monitoring." American Economic Journal: Applied Economics
		9(1):33–69.

@@

* Banerjee, A. V., Banerji, R., Duflo, E., Glennerster, R., and Khemani, S.
	2010. "Pitfalls of Participatory Programs: Evidence from a Randomized
		 Evaluation in Education in India." American Economic Journal: Economic
		 Policy, 2(1):1–30.
* Lieberman, E. S., Posner, D. N., and Tsai, L. L. 2014. "Does Information Lead
	to More Active Citizenship? Evidence from an Education Intervention in Rural
	Kenya." World Development, 60:69–83. 
* Raffler, Pia, Dan Posner, and Doug Parkerson. 2019. "The Weakness of Bottom-up
	Accountability: Experimental Evidence from the Ugandan Health Sector." URL:
	[`https://bit.ly/2L36vzI`](http://egap.org/sites/default/files/Posner%2C%20Dan_Paper.pdf)

This project re-evaluates this theoretical controversy and proposes a new field
experiment to test whether bottom-up accountability is effective to promote
better policy outcomes. The results are substantively important as well. First,
we conduct this experiment in a middle-income country, Brazil, which has
municipalities whose social indicators match both those of developed and
developing nations. Brazil is notably unequal in several characteristics, thus
allowing us to infer results for different subgroups if required. Second, we
are evaluating school constructions, and the presence of schools have essential
consequences in terms of long-run economic development, the household income
composition, and women empowerment.

@@

Interventions' dates: We ran two interventions:

@@

**Assignment Procedure**: The treatment here was assigned by the municipality
in the first intervention, and at the school level in the second and third
interventions.

@@

Then, we checked whether the one of the selected covariates is a significant
predictor of the random assignment. We presented a vector of acceptable random
assignment in these grounds, and Transparência Brasil selected the one we use
in the first intervention.

@@

1. Log of Municipal Population in 2015
2. Log of Number of Poor Families (2010 IBGE Census)
3. Log of Total Federal Transfers to the Municipality in 2016
4. IDEB Indicator for Primary School Quality (2015 Ministry of Education)
5. IDEB Indicator for Secondary School Quality (2015 Ministry of Education)

None of the pre-treatment variables were significant, as shown in Section
\ref{sec:bal-tests}.

@@

* **Treatment Groups**:
	- *Intervention 1*: Municipality with all schools funded by the Ministry of
		Education showing up in the app.
	- *Intervention 2*: School construction showing up in the app.

* **Control Groups**:
	- *Intervention 1*: Municipality with all schools funded by the Ministry of
		Education **not** showing up in the app. 
	- *Intervention 2*: Selected school constructions **not** showing up in the
		app.

* **Method of Manipulation Delivery**:
	- *Intervention 1*: The municipalities in the treatment had all their school
		constructions showing up in the app. 
	- *Intervention 2*: The schools in the treatment group were showing up in the
		app.

@@

1. Percentage of the investment executed before the impact evaluation started
	 (placebo)
2. Percentage of the investment executed by the end of the impact evaluation
	 period
3. The difference of the percentage invested in the end and in the beginning of
	 the impact evaluation period
4. Indicator for a construction finished during the impact evaluation period
5. Indicator for a construction cancelled during the impact evaluation period
6. Indicator for an updated conclusion date for the construction during the
	 impact evaluation period

@@

Below follows the CONSORT chart flow:

@@

All models use municipal-level cluster robust standard errors and state-level
fixed effects. For all models, we also run regressions without clusters or 
fixed effects.

# Other Information

@@

* **Pre-Registration**: this research has been pre-registered on the EGAP
	pre-registry tool: <https://egap.org/registration/4505>.

* **Replication Materials**: The replication materials for this project are
	available here: <https://github.com/umbertomig/tdp-accountability>.

@@

* **Conflict of Interests**: The authors of this analysis declare that there
	are no conflict of interest.

@@

We used `r R.version.string` to write the manuscript and this appendix. For the
regression models estimation, we used the `lfe` package. We employed the
`MatchIt` package to improve covariate matching. Everything in this report is
fully automated and can be reproduced using RMarkdown.

@@

# References
